---
title: "twitter-sna: tools for social network analysis with twitter data"
author: "Nicole Jess & Yuqing Liu"
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```


This GitHub repository is intended to create social network datasets based on 
data scraped from twitter and provide some useful tools for analysis.


## Obtain the Repository from GitHub
If you use Git (Torvalds et al., 2022) and have a GitHub account, either clone
or fork and clone the package to your computer using the usual Git commands
(Bryan, 2018; Bryan et al., n.d., Chapter 16). We suggest the method shown in 
[Section 16.2.2](https://happygitwithr.com/existing-github-first.html#rstudio-ide-1).

Otherwise, manually download a ZIP file from https://github.com/nrjess/twitter-sna and unzip its contents to a folder.
Either way, the files should end up in a folder called `twitter-sna` on your
computer. That folder is your local copy of the repository.

Note that the code uses relative rather than absolute folder path and
file name references. Moving or renaming subfolders and/or files may cause
problems. We have tested it only with the folder structure and file naming used
in the primary repository on GitHub.


## Understanding the Repository Structure and Contents
The structure for this repository is shown in the outline below, where folder names 
and file names are `highlighted like this` and comments are in normal text.
The repository is set up as an 
[RStudio project](https://support.rstudio.com/hc/en-us/articles/200526207-Using-RStudio-Projects). 

* `twitter-sna`: This is the root folder for the repository.
    * `data`: This folder is where the data files produced by our scripts will be 
      stored. This is a standard folder for R package structures. 
        * `Placeholder.text` This text file is just present to ensure that the
              `data` subfolder will be created when you clone the repository 
              or extract files from ZIP file copy of the repository obtained 
              from GitHub.
    * `scripts`: This folder is where you can find the key files you will need to
      use if you want to run these analyses on your own computer.
    * `.gitignore`: This file tells Git what files to ignore and omit from 
      synchronizing with the main repository on GitHub. 
    * `LICENSE`: This file contains the terms of the license that applies to all 
      source code in this repository.
    * `README.md`: This file is obtained by knitting the `README.Rmd` file and 
      is used by GitHub to display information about the package. In R Studio, you can 
      read the formatted version by opening the file and clicking the Preview 
      button.
    * `README.Rmd`: This file gives an introduction to the package. Knitting it 
      produces the `README.md` file and opens the preview automatically. 
    * `twitter-sna.Rproj`: This is an RStudio project file. It contains some 
      settings for working with the project in that software. 


## Install R Packages from CRAN
Scripts in this repository depend on having a number of R packages 
installed. Those packages are available from CRAN and can be installed by 
running the following code in the R console. 

```r 
install.packages(pkgs = c("tidyverse", "academictwitteR", "rtweet", "tidytags",
                          "textdata", "lubridate", "ergm", "networkD3", 
                          "linkcomm", "igraph", "here", "knitr", "rmarkdown"))
```

```{r install-pkgs-from-CRAN, echo = FALSE, eval = FALSE}
install.packages(pkgs = c("tidyverse", "academictwitteR", "rtweet", "tidytags",
                          "textdata", "lubridate", "ergm", "networkD3", 
                          "linkcomm", "igraph", "here", "knitr", "rmarkdown"))
```


## Access to Twitter Data
To scrape data from twitter using the method in this repository, one must have a 
Twitter developer account with a project that has Academic Research access. To 
set up such a project, follow the directions here: [Academic Research access](https://developer.twitter.com/en/products/twitter-api/academic-research). 


### Storing your Twitter access tokens

It is not safe to hard code your bearer token in your R script or have your 
bearer token in your command history.

Running the code below opens the .Renviron file and provides instructions on how
to add the bearer token, which requires the addition of a line in the .Renviron 
file, following the format: TWITTER_BEARER=YOURTOKENHERE 


```{r store-bearer-token, eval = FALSE}
library(academictwitteR)
set_bearer()
```

Save the changes to the environment and close. Go to Session and Restart R.

Now you will have a system environment variable in memory that contains your 
bearer token. You can retrieve it using the following command. 

```{r retrieve-bearer-token, eval = FALSE}
library(academictwitteR)
get_bearer()
```

## List of Tools
This section lists and briefly describes the scripts in the *twitter-sna* repository. 

- *scrape_tweets*: Create a user-defined query, scrape tweet data, and save to a 
  local repository.
- *create_social_network_data*: Restructure tweet data into a format that is 
  useful for social network analysis (edgelist and nodelist).


## Software Information
We use R Markdown to enhance reproducibility because it provides excellent 
support for generating dynamic reports (Mair, 2016). Knitting the source R 
Markdown script *`r knitr:::current_input()`* generates this Markdown file.
Knitting our other R Markdown scripts from this repository generates HTML output 
files containing explanatory text, R code, plus R output (both text and 
graphics).

- We use [RStudio](www.rstudio.org) to work with R and R markdown files. The 
  software chain that connects scripts to output files depends on which output 
  format you are using. We use the following:
    - **Rmd file > RStudio > R > rmarkdown > knitr > md file > pandoc > html file**
- We use [pandoc](https://pandoc.org) for this document. A version of pandoc 
  comes bundled with RStudio, but if you want the most recent version, download 
  it from https://pandoc.org/. 


## References
Bryan, J. (2018). Excuse me, do you have a moment to talk about version control? 
*The American Statistician, 72*(1), 20-27. 
https://doi.org/10.1080/00031305.2017.1399928

Bryan, J., The STAT 545 TAs, & Hester, J. (n.d.). 
*Happy Git and GitHub for the useR*. https://happygitwithr.com

Mair, P. (2016). Thou shalt be reproducible! A technology perspective. Frontiers
in Psychology, 7(1079), 1-17. http://dx.doi.org/10.3389/fpsyg.2016.01079

Torvalds, L., Hamano, J. C., & other contributors to the Git Project. (2022). 
Git for Windows (Version 2.37.1) [Computer program]. Brooklyn, NY: Software 
Freedom Conservancy. Retrieved from https://git-scm.com
