---
title: "Creating Social Network Datasets"
author: "Nicole Jess, Yuqing Liu"
date: "`r format(Sys.time(), '%Y-%m-%d, %H:%M:%S %Z')`"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding:  hide
params: 
  outfile: "create-social-network-data.html"
---

# Purpose
The script restructures data scraped from twitter into a format that can be used for social network analysis.

```{r global-options, include=FALSE}
# Create a custom chunk hook/option for controlling font size in chunk & output.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$cfsize != "normalsize", paste0("\n \\", options$cfsize,"\n\n", 
                                              x, "\n\n \\normalsize"), x)
  })

# Global chunk options (over-ridden by local chunk options)
knitr::opts_chunk$set(include  = TRUE, echo = TRUE, error = TRUE, 
                      message = FALSE, warning = FALSE, 
                      cfsize = "footnotesize", fig.height=8)

# Declare location of this script relative to the project root directory.
here::i_am(path = "scripts/create_social_network_data.Rmd")
```

# Load R Packages
Load contributed R packages that we need to get additional functions. 

``` {r load-packages}
library(tidyverse)          # dplyr, ggplot2, tidyr, etc.
library(here)               # for here()
library(academictwitteR)    # for get_user_profile()
library(linkcomm)           # for integer.edgelist()
library(igraph)

```

# Load data

``` {r load-data}
# Data file created by "scripts/scrape_tweets.Rmd"
load(file=here("data/t2.RData"))
load(file=here("data/tweet_authors.RData"))
```

# Edgelist

## Replys
```{r reply-edges}
replies <- t2 %>%
  select(id_str, created_at, in_reply_to_user_id_str, full_text) %>%
  drop_na() %>%
  mutate(edge_type="reply") %>%
  left_join(tweet_authors, by=c("id_str"="id")) %>%
  rename(tweet_id=id_str,
         from=author_id,
         to=in_reply_to_user_id_str,
         text=full_text) %>%
  relocate(tweet_id, created_at, from, to, text, edge_type)

```

## Retweets
```{r retweet-edges}

retweets<- t2$retweeted_status %>%
  map("user") %>%
  map_df(bind_rows) %>%
  mutate(tweet_id=t2$id_str,
         text=t2$full_text,
         created_at=t2$created_at,
         edge_type="retweets") %>%
  drop_na(id_str) %>%
  left_join(tweet_authors, by=c("tweet_id"="id")) %>%
  rename(from=author_id,
         to=id_str) %>%
  select(tweet_id, created_at, from, to, text, edge_type)


```

## Quotes
```{r quote-edges}

quotes<- t2$quoted_status %>%
  map("user") %>%
  map_df(bind_rows) %>%
  mutate(tweet_id=t2$id_str,
         text=t2$full_text,
         created_at=t2$created_at,
         edge_type="quotes") %>%
  drop_na(id_str) %>%
  left_join(tweet_authors, by=c("tweet_id"="id")) %>%
  rename(from=author_id,
         to=id_str) %>%
  select(tweet_id, created_at, from, to, text, edge_type)

```

## Mentions
```{r mention-edges}

mentions <- t2$entities %>%
  map("user_mentions") %>%
  map("id_str") %>%
  plyr::ldply(rbind) %>%
  mutate(tweet_id=t2$id_str,
         text=t2$full_text,
         created_at=t2$created_at,
         edge_type="mentions") %>%
  pivot_longer(-c("tweet_id","text","created_at","edge_type"), values_to="to") %>%
  drop_na(to) %>%
  left_join(tweet_authors, by=c("tweet_id"="id")) %>%
  rename(from=author_id) %>%
  select(tweet_id, created_at, from, to, text, edge_type)  

```

## Combine edge types

You can remove edge types that you are not interested in. For example, if you do not want to include replies, delete "replies" from the list below.

```{r all-edges}

edgelist <- bind_rows(replies,
                      retweets,
                      quotes,
                      mentions)

```

# Integer edgelist

```{r integer-edges}

# integer edgelist
network1 <- edgelist %>%
  select(from,to) %>%
  as.matrix() %>%
  integer.edgelist()

edges <- network1$edges %>%
  as.data.frame() %>%
  rename(from = V1,
         to = V2) %>%
  group_by(from,to) %>%
  summarize(weight = n()) %>%
  filter(from != to) %>%    # remove self-ties
  as.data.frame()

# save indegree and outdegree to use as node attributes
indegree <- edges %>%
  group_by(to) %>%
  summarize(indegree=n()) %>%
  as.data.frame()

outdegree <- edges %>%
  group_by(from) %>%
  summarize(outdegree=n()) %>%
  as.data.frame()

```


# Nodelist

## Scrape profiles
```{r profile-scrape}
nodelist <- c(edgelist$from, edgelist$to) %>%
  unique() %>%
  get_user_profile(get_bearer()) %>%
  select(-starts_with("entities"))

# integer IDs to node attributes
nodes <- network1$nodes %>%
  as.data.frame() %>%
  rename(nodes = 1) %>%
  mutate(id = rownames(.)) %>%
  left_join(nodelist, by="id") %>%
  left_join(indegree, by=c("nodes"="to")) %>%
  left_join(outdegree, by=c("nodes"="from")) %>%
  mutate_at(vars(c(indegree,outdegree)), ~replace(., is.na(.), 0)) %>%
  mutate(degree = indegree + outdegree)

# create an igraph object
net1 <- graph_from_data_frame(d = edges,
                              vertices = nodes,
                              directed = TRUE)

plot(net1)

```


# Save Data

## T2 edges and nodes
```{r save-t2-data}
# here() creates a path relative to the location of the twitter-sna.proj file
# that will be portable across local repositories on different computers.

save(edgelist, file=here("data/twitter_edges.RData"))
save(nodelist, file=here("data/twitter_nodes.RData"))
save(edges, file=here("data/int_edges.RData"))
save(nodes, file=here("data/int_nodes.RData"))

```
