---
title: "Creating Social Network Datasets"
author: "Nicole Jess, Yuqing Liu"
date: "`r format(Sys.time(), '%Y-%m-%d, %H:%M:%S %Z')`"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding:  show
params: 
  outfile: "create-social-network-data.html"
---

# Purpose
The script restructures data scraped from twitter into a format that can be used for social network analysis.

```{r global-options, include=FALSE}
# Create a custom chunk hook/option for controlling font size in chunk & output.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$cfsize != "normalsize", paste0("\n \\", options$cfsize,"\n\n", 
                                              x, "\n\n \\normalsize"), x)
  })

# Global chunk options (over-ridden by local chunk options)
knitr::opts_chunk$set(include  = TRUE, echo = TRUE, error = TRUE, 
                      message = FALSE, warning = FALSE, 
                      cfsize = "footnotesize", fig.height=8)

# Declare location of this script relative to the project root directory.
here::i_am(path = "scripts/create_social_network_data.Rmd")
```

# Load R Packages
Load contributed R packages that we need to get additional functions. 

``` {r load-packages}
library(tidyverse)          # dplyr, ggplot2, tidyr, etc.
library(here)               # for here()
library(knitr)              # for kable()
library(academictwitteR)    # for get_user_profile()
library(linkcomm)           # for integer.edgelist()
library(igraph)             # for graph_from_data_frame()

```

# Load data

``` {r load-data}
# Data file created by "scripts/scrape_tweets.Rmd"
load(file=here("data/t2.RData"))
load(file=here("data/tweet_authors.RData"))
```

# Edgelist

## Replys

```{r reply-edges}
replies <- t2 %>%
  select(id_str, created_at, in_reply_to_user_id_str, full_text) %>%
  drop_na() %>%
  mutate(edge_type="replies") %>%
  left_join(tweet_authors, by=c("id_str"="id")) %>%
  rename(tweet_id=id_str,
         from=author_id,
         to=in_reply_to_user_id_str,
         text=full_text) %>%
  relocate(tweet_id, created_at, from, to, text, edge_type)

```

## Retweets
```{r retweet-edges}

retweets<- t2$retweeted_status %>%
  map("user") %>%
  map_df(bind_rows) %>%
  mutate(tweet_id=t2$id_str,
         text=t2$full_text,
         created_at=t2$created_at,
         edge_type="retweets") %>%
  drop_na(id_str) %>%
  left_join(tweet_authors, by=c("tweet_id"="id")) %>%
  rename(from=author_id,
         to=id_str) %>%
  select(tweet_id, created_at, from, to, text, edge_type)


```

## Quotes
```{r quote-edges}

quotes<- t2$quoted_status %>%
  map("user") %>%
  map_df(bind_rows) %>%
  mutate(tweet_id=t2$id_str,
         text=t2$full_text,
         created_at=t2$created_at,
         edge_type="quotes") %>%
  drop_na(id_str) %>%
  left_join(tweet_authors, by=c("tweet_id"="id")) %>%
  rename(from=author_id,
         to=id_str) %>%
  select(tweet_id, created_at, from, to, text, edge_type)

```


## Mentions
```{r mention-edges}

mentions <- t2$entities %>%
  map("user_mentions") %>%
  map("id_str") %>%
  plyr::ldply(rbind) %>%
  mutate(tweet_id=t2$id_str,
         text=t2$full_text,
         created_at=t2$created_at,
         edge_type="mentions") %>%
  pivot_longer(-c("tweet_id","text","created_at","edge_type"), values_to="to") %>%
  drop_na(to) %>%
  left_join(tweet_authors, by=c("tweet_id"="id")) %>%
  rename(from=author_id) %>%
  select(tweet_id, created_at, from, to, text, edge_type)  

```

## Combine edge types

You can remove edge types that are not of interest for your project. For example, if you do not want replies to be counted as edges in your network, delete "replies" from the code below.

```{r all-edges}

# combine all edge types, exclude any edge types that you do not want to include in your edgelist
edgelist <- bind_rows(replies,
                      retweets,
                      quotes,
                      mentions)

# drop edges with missing data in 'from' or 'to'
edgelist <- edgelist %>%
  drop_na(from,to)

```

Altogether, this gives us a total of `r nrow(edgelist)` edges in this network. Table \@ref(tab:edges-table) shows the total number of each edge type in this network.

```{r edges-table}

TC <- paste0("Total edges by edge type ")

table(edgelist$edge_type) %>%
  addmargins(1,sum) %>%
  kable(col.names=c("edge type","n"),
        caption=TC)


```

# Integer edgelist

Some of the tools we work with require that the edge IDs are integers starting with 1. This code creates and edgelist that meets these requirements.

```{r integer-edges}

# integer edgelist
network1 <- edgelist %>%
  select(from,to) %>%
  as.matrix() %>%
  integer.edgelist()

int_edges <- network1$edges %>%
  as.data.frame() %>%
  rename(from = V1,
         to = V2) %>%
  group_by(from,to) %>%
  summarize(weight = n()) %>%
  as.data.frame()

# save indegree and outdegree to use as node attributes
indegree <- int_edges %>%
  group_by(to) %>%
  summarize(indegree=n()) %>%
  as.data.frame()

outdegree <- int_edges %>%
  group_by(from) %>%
  summarize(outdegree=n()) %>%
  as.data.frame()

```


# Nodelist

This code creates a list of all the nodes in the network and scrapes data from the Twitter user profiles. 

```{r profile-scrape}
nodelist <- c(edgelist$from, edgelist$to) %>%
  unique() %>%
  get_user_profile(get_bearer()) %>%
  select(-starts_with("entities"))

# match integer IDs to node attributes
int_nodes <- network1$nodes %>%
  as.data.frame() %>%
  rename(nodes = 1) %>%
  mutate(id = rownames(.)) %>%
  left_join(nodelist, by="id") %>%
  left_join(indegree, by=c("nodes"="to")) %>%
  left_join(outdegree, by=c("nodes"="from")) %>%
  mutate_at(vars(c(indegree,outdegree)), ~replace(., is.na(.), 0)) %>%
  mutate(degree = indegree + outdegree)


```


# Testing an *igraph* object

This code creates and plots an *igraph* object to test that the integer edgelist works properly. Figure \@ref(fig:igraph) probably doesn't look very nice but for this script we just want to confirm that we are able to make a graph. Other scripts in this repository have code for making better graphs.

```{r igraph, fig.width=7, fig.height=4, fig.cap=FC}

FC <- paste0("\\label{fig:igraph}",
            "Plot of network ")

# create an igraph object
net1 <- graph_from_data_frame(d = int_edges,
                              vertices = int_nodes,
                              directed = TRUE)

plot(net1)

```


# Save Data

```{r save-t2-data}
# here() creates a path relative to the location of the twitter-sna.proj file
# that will be portable across local repositories on different computers.

save(edgelist, file=here("data/edgelist.RData"))
save(nodelist, file=here("data/nodelist.RData"))
save(int_edges, file=here("data/int_edges.RData"))
save(int_nodes, file=here("data/int_nodes.RData"))

```
