---
title: "Path Visualizations"
author: "Nicole Jess, Hasan Gokberk Bayhan, Yuqing Liu"
date: "`r format(Sys.time(), '%Y-%m-%d, %H:%M:%S %Z')`"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding:  hide
params: 
  outfile: "path_visualizations.html"
---

# Purpose
The script creates a dynamic visualization of a network based on an edgelist created from Twitter data.

```{r global-options, include=FALSE}
# Create a custom chunk hook/option for controlling font size in chunk & output.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$cfsize != "normalsize", paste0("\n \\", options$cfsize,"\n\n", 
                                              x, "\n\n \\normalsize"), x)
  })

# Global chunk options (over-ridden by local chunk options)
knitr::opts_chunk$set(include  = TRUE, echo = TRUE, error = TRUE, 
                      message = FALSE, warning = FALSE, 
                      cfsize = "footnotesize", fig.height=8)

# Declare location of this script relative to the project root directory.
here::i_am(path = "scripts/path_visualizations.Rmd")
```

# Load R Packages
Load contributed R packages that we need to get additional functions. 

``` {r load-packages}
library(tidyverse)          # dplyr, ggplot2, tidyr, etc.
library(here)               # for here()
library(lubridate)          # for floor_date()
library(linkcomm)           # for integer.edgelist()
library(networkDynamic)     # for networkDynamic()
library(ndtv)               # for compute.animation(), render.d3movie(), transmissionTimeline()
library(tsna)               # for tPath()

```

# Load data

``` {r load-data}
# Data file created by "scripts/create_social_network_data.Rmd"
load(file=here("data/edgelist.RData"))
load(file=here("data/int_nodes.RData"))
```

# Choose a duration

The duration defines how long tweets will remain visible in the dynamic visualization. For this example, we chose to allow the effect of a tweet to last 7 days.

```{r }

duration = 30

# define onset and terminus
start <- edgelist$created_at %>% min() %>% as.Date() %>% floor_date(.,unit="day") # first date observed in edgelist

edgelist$onset  <- difftime(edgelist$created_at, start, units="days") %>%
  as.double() %>% 
  floor()

edgelist$terminus  <- edgelist$onset + duration  # can change--we chose to have the effect of a tweet last for 7 days

```

# Formatting the edgelist

This does the following:
* applies integer IDs to edges
* formats dates correctly
* simplifies variable names

```{r}

# save integer ids
node_ids <- int_nodes %>% select(nodes, id, username)

# add integer ids to edgelist
edges <- edgelist %>% 
  left_join(node_ids, by=c('from'='id')) %>%  # join with integer node ids for 'from'
  select(-from) %>%
  rename(head = nodes,
         from = username) %>%
  left_join(node_ids, by=c('to'='id')) %>%  # join with integer node ids for 'to'
  select(-to) %>%
  rename(tail = nodes,
         to = username) %>%
  filter(head != tail) %>%  # remove self-ties
  relocate(from,to)

# created_at should be formatted as date
edges$created_at <- as.Date(edges$created_at, format="%Y-%m-%d")

# tail and head should be formatted as numeric
edges$tail <- as.numeric(edges$tail)
edges$head <- as.numeric(edges$head)

# give follower count a better name
int_nodes$follower_count <- int_nodes$public_metrics$followers_count 

```

# Combine overlapping edges

Edges that overlap need to be combined. For example, if node A and node B have an edge that lasts from time 1-7 and another edge that lasts from time 3-9, this would be combined into a single edge that lasts from time 1-9.

```{r combine-edges}

es <- edges %>%
  arrange(tail,head,onset) %>% # sort
  group_by(tail,head) %>%
  mutate(index = cumsum(cummax(lag(terminus, default = first(terminus))) < onset)) %>%  # index for distinct time intervals
  group_by(tail,head,from,to,index) %>%  # combine rows that represent overlapping intervals between the same 2 nodes
  summarise(onset = min(onset),
            terminus = max(terminus),  
            edge_type = toString(unique(edge_type)),  # lists for multiple edge_types (ex- retweet, mention)
            text = toString(unique(text))) %>%   # list text of multiple tweets where applicable
  select(onset,terminus,tail,head,from,to,text,edge_type) %>%
  as.data.frame()

```

# Formatting the nodelist

This adds onset and terminus information to the nodelist which allows node characteristics to be dynamic. 

```{r prep-nodes}

# format node attributes for dynamic network
vs <- rbind(es %>%   # extract all 'tail' nodes from edgelist
              select(onset,terminus,tail) %>% 
              rename(vertex.id=tail),
            es %>%   # extract all 'head' nodes from edgelist
              select(onset,terminus,head) %>% 
              rename(vertex.id=head)) %>%  
  distinct(onset, terminus, vertex.id) %>%  # keep only unique combinations
  left_join(int_nodes, by=c('vertex.id'='nodes')) %>%  #join with node attributes
  select(onset,terminus,vertex.id,username,description,location,verified,follower_count,degree) %>%
  as.data.frame() %>%
  arrange(vertex.id) 

```

# Create a dynamic network object

Includes the edge and node attributes needed for the visualization.

```{r}

# create dynamic network (for full network)
net1.dyn <- networkDynamic(edge.spells=es, 
                           vertex.spells=vs, 
                           create.TEAs=TRUE,  # create dynamic attributes
                           edge.TEA.names=c('from','to','text','edge_type'),  # list dynamic edge attributes in order starting w/ col 5 of es
                           vertex.TEA.names=c('username','description','location','verified','follower_count','degree')) # list dynamic node attributes in order starting w/ col 4 of vs 

```


# Path Visualizations

```{r}

actor = c(33) # 33 is the integer ID for DeAngelisCorey

# extract the forward reachable paths from a node
actor_fwd <- tPath(net1.dyn,
                   v = actor,
                   direction = "fwd",
                   start = 0,
                   end = max(es$terminus))

```

## Static visualizations

```{r fwd-plot1, eval=FALSE}
plot(actor_fwd)

```

```{r fwd-plot2, error=FALSE}
plot(transmissionTimeline(actor_fwd,
                     jitter=TRUE,
                     main='Paths of @DeAngelisCorey'))
```

```{r fwd-plot3, eval=FALSE}
plotPaths(net1.dyn,
          actor_fwd,
          displaylabels=FALSE,
          vertex.col="white",
          edge.label.col = rgb(0,0,0,0))
```

## Dynamic Visualization

```{r dyn-fwd-plot-prep}

# pull out the vertex.ids and tdist (onset must be greater than this)
actor <- as.data.frame(actor_fwd) %>%
  mutate(vertex.id = row_number()) %>%
  filter(tdist != 'Inf') %>%
  select(vertex.id, tdist)

# tail must be an element of actor$vertex.id and
# the corresponding onset must be greater or equal to the tdist associated with that vertex.id
es.actor <- es %>%
  right_join(actor, by = c("tail" = "vertex.id")) %>%
  filter(onset>=tdist)

# filter nodes--keep only those in edgelist
vs2 <- vs %>%
  filter(vertex.id %in% es.actor$tail | vertex.id %in% es.actor$head) %>%
  mutate(vertex.id = as.numeric(factor(username)),
         att = ifelse(onset<10,0,1))

# assign new node ids (must be inceasing start with 1)
node_ids2 <- vs2 %>% select(vertex.id, username) %>% unique()

# match edgelist with new node ids
es2 <- es.actor %>%
  select(-head, -tail) %>%
  left_join(node_ids2, by=c('from'='username')) %>%
  rename(head = vertex.id) %>%
  left_join(node_ids2, by=c('to'='username')) %>%
  rename(tail = vertex.id) %>%
  filter(head %in% node_ids2$vertex.id & tail %in% node_ids2$vertex.id) %>%
  select(onset,terminus,tail,head,from,to,text,edge_type)

```

```{r dyn-fwd-network}

# create dynamic network
net2.dyn <- networkDynamic(edge.spells=es2,
                           vertex.spells=vs2,
                           create.TEAs=TRUE,  # create dynamic attributes
                           edge.TEA.names=c('from','to','text','edge_type'),  # list dynamic edge attributes in order starting w/ col 5 of es
                           vertex.TEA.names=c('username','description','location','verified','follower_count','indegree','att')) # list dynamic node attributes in order starting w/ col 4 of vs


# length of these edgelists needs to match
nrow(es2)
nrow(as.data.frame(net2.dyn))

```


```{r dyn-fwd-visualization}

# set time interval
net2.dyn <- compute.animation(net2.dyn,
                              slice.par=list(start=min(es2$onset),
                                             end=max(es2$terminus),
                                             interval=1,  # currently daily, set to 7 for weekly
                                             aggregate.dur=1,  # currently daily, set to 7 for weekly
                                             rule="earliest"))

# dynamic visualization
render.d3movie(net2.dyn,
               usearrows = FALSE,
               edge.lwd = 2,
               edge.col = 'grey',
               edge.tooltip = function(slice) {     # display edge attributes when you click on an edge
                 paste("<b>From:</b>", (slice %e% "from"),      "<br>",
                       "<b>To:</b>",   (slice %e% "to"),        "<br>",
                       "<b>Type:</b>", (slice %e% "edge_type"), "<br>",
                       "<b>Text:</b>", (slice %e% "text"))
                 },
               vertex.cex = .3, # add sizing by attrbute
                    # vertex.cex = .3, # add sizing by attrbute
               vertex.col =  function(slice) {
                 case_when((slice %v% "verified" == TRUE) ~ 'green',  # verified nodes are green, all others are grey
                           (slice %v% "verified" == FALSE) ~ 'grey',
                            is.na(slice %v% "verified") ~ 'grey')
                 },
               vertex.sides =  function(slice) {
                 case_when((slice %v% "username" == "DeAngelisCorey") ~ 3,   # DeAngelisCorey is a triangle
                           (slice %v% "username" != "DeAngelisCorey") ~ 50)  # everyone else is a circle
                 },  # change node shapes by attribute
               vertex.tooltip = function(slice) {     # display node attributes when you click on an node
                 paste("<b>Username:</b>",    (slice %v% "username") ,    "<br>",
                       "<b>Location:</b>",    (slice %v% "location") ,    "<br>",
                       "<b>Description:</b>", (slice %v% "description") , "<br>",
                       "<b>Followers:</b>",   (slice %v% "follower_count"))
                 },
                 #displaylabels = TRUE,
                 #label = net1.dyn %v% "username",
                 #label.cex = .6,
                 #label.col = "blue",
               main = 'school choice ties: @DeAngelisCorey forward paths',
               output.mode = 'htmlWidget')

```

