---
title: "Scraping Tweets"
author: "Nicole Jess, Yuqing Liu"
date: "`r format(Sys.time(), '%Y-%m-%d, %H:%M:%S %Z')`"
output: 
  bookdown::html_document2:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding:  hide
params: 
  outfile: "scrape_tweets.html"
---

# Purpose
The script scrapes Twitter based on parameters defined by the user and saves a 
copy to an .RData file stored in the local repository where we can easily access 
it from other scripts.

```{r global-options, include=FALSE}
# Create a custom chunk hook/option for controlling font size in chunk & output.
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$cfsize != "normalsize", paste0("\n \\", options$cfsize,"\n\n", 
                                              x, "\n\n \\normalsize"), x)
  })

# Global chunk options (over-ridden by local chunk options)
knitr::opts_chunk$set(include  = TRUE, echo = TRUE, error = TRUE, 
                      message = FALSE, warning = FALSE, 
                      cfsize = "footnotesize", fig.height=8)

# Declare location of this script relative to the project root directory.
here::i_am(path = "scripts/scrape_tweets.Rmd")
```

# Load R Packages
Load contributed R packages that we need to get additional functions. 

``` {r load-packages}
library(tidyverse)          # dplyr, ggplot2, tidyr, etc.
library(here)               # for here()
library(academictwitteR)    # for build_query(), count_all_tweets(), get_all_tweets(), get_bearer()
library(rtweet)             # for rtweet_app(), auth_as(), lookup_tweets()
library(knitr)              # for kable()
library(lubridate)          # for floor_date()
```


# Define your query

Multiple keywords can be used as shown:

* query = c("cat", "dog") returns tweets that contain the word "cat" or the word "dog"
* query = c("cat dog") returns tweets that contain both words 

The query can also be restricted by users, exclude keywords, and filter by tweet 
types. Full list of options shown here: [build_query](https://rdrr.io/cran/academictwitteR/man/build_query.html)

``` {r query}
# define hashtags/keywords
# "," means "or"
# " " means "and"
query1 <- build_query(query = c("school vouchers", "#schoolchoice"))

# define time intervals
# time interval used to measure 'before' sentiment
time_interval1 <- data.frame(start = "2022-04-01T00:00:00Z",
                             end = "2022-04-11T00:00:00Z")
# time interval used to create edgelist
time_interval2 <- data.frame(start = "2022-04-11T00:00:00Z",
                             end = "2022-04-21T00:00:00Z")
# time interval used to measure 'after' sentiment
time_interval3 <- data.frame(start = "2022-04-21T00:00:00Z",
                             end = "2022-05-01T00:00:00Z")
```


## Check tweet counts

```{r total-tweets}
# how many tweets will this query yield?
count.tweets <- count_all_tweets(query = c(query1),
                                 start_tweets=time_interval1$start,
                                 end_tweets=time_interval3$end,
                                 get_bearer(),
                                 n=Inf,
                                 verbose=FALSE)

```

The total number of tweets scraped in this query is `r sum(count.tweets$tweet_count)`.

Table \@ref(tab:tweets-by-interval) shows the number of tweets that fall into each of the time intervals and Figure \@ref(fig:timeseries-plot) shows the number of tweets plotted over time.

```{r tweets-by-interval}
tab <- count.tweets %>%
  mutate(time_interval=case_when((start < time_interval1$end)~1,
                                 (start >= time_interval2$start & start < time_interval2$end)~2,
                                 (start >= time_interval3$start)~3)) %>%
  group_by(time_interval) %>%
  summarize(tweet_count=sum(tweet_count))

TC <- paste0("Number of tweets by time interval ")

kable(tab,
      col.names=c("interval","number of tweets"),
      caption=TC)
```

```{r timeseries-plot, fig.width=7, fig.height=4, fig.cap=FC}

# Figure caption
FC <- paste0("\\label{fig:timeseries-plot}",
            "Timeseries plot of tweets over time ")

ggplot(data = count.tweets,
       aes(x=as.Date(end), y=tweet_count, group=1)) +
  geom_line() +
  geom_vline(xintercept=c(as.Date(time_interval2$start),
                          as.Date(time_interval3$start)),
             linetype=2,
             color="red") +
  labs(x="Date", y="number of tweets") +
  scale_x_date(date_breaks = "1 week",
               date_labels = "%m-%d-%Y") +
  theme_classic()
```


# Scrape tweets
```{r tweet-scrape}
tweets <- get_all_tweets(query = c(query1),
                         start_tweets=time_interval1$start,
                         end_tweets=time_interval3$end,
                         get_bearer(),
                         n=Inf) # default is n=100 (pulls 100 most recent tweets), set to n=Inf to get all the tweets

# auth for rtweet package using stored bearer token
auth <- rtweet_app(bearer_token = get_bearer())

tweets_processed <- lookup_tweets(tweets$id,
                                  retryonratelimit=TRUE,
                                  token=auth_as(auth))

```

```{r split-time-intervals}
t1 <- tweets_processed %>%
  filter((created_at > time_interval1$start) & (created_at <= time_interval1$end))

t2 <- tweets_processed %>%
  filter((created_at > time_interval2$start) & (created_at <= time_interval2$end))

t3 <- tweets_processed %>%
  filter((created_at > time_interval3$start) & (created_at <= time_interval3$end))
```


## Keyword usage

Figure \@ref(fig:keyword-plot) shows the number of tweets plotted over time.

```{r keyword-plot, fig.width=7, fig.height=4, fig.cap=FC}
tab1 <- tweets_processed %>%
  mutate(day = floor_date(created_at, "day"),
         `#charterschool`= if_else(grepl("charter school", text), 1, 0),
         `#schoolchoice`= if_else(grepl("#schoolchoice", text), 1, 0)) %>%
  pivot_longer(cols=starts_with("#"),
               names_to = "keyword",
               values_to = "keyword_used",
               values_drop_na = TRUE) %>%
  group_by(day, keyword) %>%
  summarize(count = sum(keyword_used))

FC <- paste0("\\label{fig:keyword-plot}",
            "Use of keywords over time ")

ggplot(data = tab1,
       aes(x=as.Date(day), y=count, color=keyword)) +
  geom_line() +
  geom_vline(xintercept=c(as.Date(time_interval2$start),
                          as.Date(time_interval3$start)),
             linetype=2,
             color="red") +
  labs(title="Use of keywords over time", x="Date",y="number of tweets") +
  scale_x_date(date_breaks = "6 days",
               date_labels = "%m-%d-%Y") +
  theme_classic() +
  theme(legend.title = element_blank())
```


``` {r tweet_attributes}
tweet_authors <- tweets %>%
  select(id, author_id)

```

# Save Data

```{r save-data}
# here() creates a path relative to the location of the twitter-sna.proj file
# that will be portable across local repositories on different computers.

save(t1, file=here("data/t1.RData"))
save(t2, file=here("data/t2.RData"))
save(t3, file=here("data/t3.RData"))
save(tweet_authors, file=here("data/tweet_authors.RData"))

```

